---
title: 'Reproducible Research: Peer Assessment 1'
date: "Febrero 2019"
output:
  html_document: default
  pdf_document: default
---

# Data

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA).
Número de pasos que se toman en un intervalo de 5 minutos.

date: The date on which the measurement was taken in YYYY-MM-DD format.

interval: Identifier for the 5-minute interval in which measurement was taken.
Identificador del intervalo de 5 minutos en el que se realizó la medición.

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

```{r, comment="", warning=F, message=F, echo=F}
rm(list = ls())

# Library

library(ggplot2)
library(plyr)
library(lattice)

# Loading and preprocessing the data

activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
pasos <- with(activity, tapply(steps, date, sum))
steps_day <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
steps_interval <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
```

## What is mean total number of steps taken per day?
¿Cuál es el número total medio de pasos tomados por día?

```{r, comment="", warning=F, message=F, echo=F}
hist(steps_day$steps,
     xlab = "Steps taken Daily",
     ylab = "Frequency",
     main = "Total Number of Steps Taken per Day",
     col = "pink")

dev.copy(png, file = "plot.1.png", height = 480, width = 480)
dev.off()

summary(steps_day$steps)

mean <- mean(steps_day$steps, na.rm = TRUE)
median <- median(steps_day$steps, na.rm = TRUE)

mean
median
```

## What is the average daily activity pattern?
¿Cuál es el patrón de actividad diaria promedio?

```{r, comment="", warning=F, message=F, echo=F}
str(activity)

plot(steps ~ interval, data = steps_interval,
     type="l", 
     lwd=2, 
     xlab="Time", 
     ylab="Number of steps", 
     main="Average Daily Activity Pattern")

dev.copy(png, file = "plot.2.png", height = 480, width = 480)
dev.off() 
```

## Imputing missing values
Imputando valores perdidos

```{r, comment="", warning=F, message=F, echo=F}
value_na <- !complete.cases(activity)
num_na <- sum(is.na(activity))
num_na

data <- activity[value_na == TRUE, ]
data[, 1] <- steps_interval$steps
data2 <- rbind(activity[complete.cases(activity), ], data)
steps_day2 <- aggregate(steps ~ date, data = data2, sum)

hist(steps_day2$steps,
     xlab = "Day",
     ylab = "Frequency",
     main = "Total Steps per Day",
     col = "light blue")

dev.copy(png, file = "plot.3.png", height = 480, width = 480)
dev.off() 

summary(steps_day2)

mean <- mean(steps_day2$steps, na.rm = TRUE)
median <- median(steps_day2$steps, na.rm = TRUE)

mean
median
```

## Are there differences in activity patterns between weekdays and weekends?

```{r, comment="", warning=F, message=F, echo=F}

weekdays <- weekdays(data2$date)
data2$weekdays <- ifelse(weekdays == "domingo" | weekdays == "sábado", 
    "Weekend", "Weekday")

str(data2)

steps_weekd <- aggregate(data2$steps, by = list(data2$interval, data2$weekdays), mean)
names(steps_weekd) <- c("interval", "weekdays", "steps")

xyplot(steps ~ interval | weekdays, steps_weekd,
       type = "l",
       layout = c(1, 2),
       xlab = "Interval",
       ylab = "Number of steps",
       main = "Weekends and Weekdays")

dev.copy(png, file = "plot.4.png", height = 480, width = 480)
dev.off() 
```
